Pleiotropy of expression quantitative trait loci increases with the number of regulated genes, active cell types and tissues and bound transcription factors

### 2022-11-27 ms/ms03_results.tex
ENH

\subsection*{A pipeline to systematically colocalize eQTLs and GWAS variants}

To better understand the role of gene regulatory variants for variant pleiotropy, I developed a pipeline to systematically colocalize eQTLs and GWAS variants (Section Methods).
%
127 eQTL studies were downloaded from the EBI eQTL catalogue, which aims to provide uniformly processed eQTLs in many tissues and cell types \citep{2021.Alasoo.Kerimov}.

@@

\subsection*{Do pleiotropic eQTLs have stronger effect sizes?}

Pleiotropic eQTLs affect simultaneously more trait classes.
%
Therefore, I wonder about the relationship between pleiotropy, effect size and significance of eQTLs.
%
I found that the mean of the absolute eQTL effect sizes (beta) decreased between GWAS class counts 1 and 5 (Fig. \ref{fig:beta_pval}a).
%
Regarding the eQTL significance (Negative decimal logarithm of the p-values), I found decreasing mean values for GWAS class counts between 1 and 5 (Fig. \ref{fig:beta_pval}b).

Then I carried out the same analysis for the GWAS effect size (beta) and significance (p-value).
%
I found that the mean of the absolute GWAS effect sizes (beta) decreased between GWAS class counts 1 and 5 (Fig. \ref{fig:beta_pval}b).
%
Regarding the GWAS variant significance (Negative decimal logarithm of the p-values), I found increasing mean values for GWAS class counts between 1 and 5 (Fig. \ref{fig:beta_pval}b).

These observations suggest that the strength of both eQTL and GWAS effects decrease with the pleiotropy (Fig. \ref{fig:beta_pval}a,b).

>>>>

\subsection*{Systematic colocalization of variants from genome-wide associations studies and expression quantitative trait loci}

To better understand the role of gene regulatory variants for variant pleiotropy,
I developed a pipeline to systematically colocalize variants from genome-wide associations
studies (GWAS) and expression quantitative trait loci (eQTL) (Section Methods).
%
127 eQTL studies were downloaded from the EBI eQTL catalogue, which aims to provide uniformly processed eQTLs in many tissues and cell types \citep{2021.Alasoo.Kerimov}.

@@

%\subsection*{Do pleiotropic eQTLs have stronger effect sizes?}
%
%Pleiotropic eQTLs affect simultaneously more trait classes.
%%
%Therefore, I wonder about the relationship between pleiotropy, effect size and significance of eQTLs.
%%
%I found that the mean of the absolute eQTL effect sizes (beta) decreased between GWAS class counts 1 and 5 (Fig. \ref{fig:beta_pval}a).
%%
%Regarding the eQTL significance (Negative decimal logarithm of the p-values), I found decreasing mean values for GWAS class counts between 1 and 5 (Fig. \ref{fig:beta_pval}b).
%
%Then I carried out the same analysis for the GWAS effect size (beta) and significance (p-value).
%%
%I found that the mean of the absolute GWAS effect sizes (beta) decreased between GWAS class counts 1 and 5 (Fig. \ref{fig:beta_pval}b).
%%
%Regarding the GWAS variant significance (Negative decimal logarithm of the p-values), I found increasing mean values for GWAS class counts between 1 and 5 (Fig. \ref{fig:beta_pval}b).
%
%These observations suggest that the strength of both eQTL and GWAS effects decrease with the pleiotropy (Fig. \ref{fig:beta_pval}a,b).

### 2022-11-27 ms/ms06_floats.tex
ENH

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Fig 8: beta and log pval
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\begin{figure}[!tbp]
\centering
%
\begin{subfigure}[]{.33\textwidth}
\textbf{a}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta.py/eqtl_beta.png}
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{b}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta.py/gwas_beta.png}
\end{subfigure}

\begin{subfigure}[]{.33\textwidth}
\textbf{c}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_logpval.py/eqtl.png}
\end{subfigure}
%
\begin{subfigure}[]{.33\textwidth}
\textbf{d}
\\
\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_logpval.py/gwas.png}
\end{subfigure}

\caption{eQTL and GWAS effect size/beta (\textbf{a,b}) and eQTL and GWAS significance/p-value (\textbf{c,d}) as a function of the GWAS class count.} \label{fig:beta_pval}
%
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Fig 8: beta and log pval
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\begin{figure}[!tbp]
%\centering
%%
%\begin{subfigure}[]{.33\textwidth}
%\textbf{a}
%\\
%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta.py/eqtl_beta.png}
%\end{subfigure}
%%
%\begin{subfigure}[]{.33\textwidth}
%\textbf{b}
%\\
%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta.py/gwas_beta.png}
%\end{subfigure}
%
%\begin{subfigure}[]{.33\textwidth}
%\textbf{c}
%\\
%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_logpval.py/eqtl.png}
%\end{subfigure}
%%
%\begin{subfigure}[]{.33\textwidth}
%\textbf{d}
%\\
%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_logpval.py/gwas.png}
%\end{subfigure}
%
%\caption{eQTL and GWAS effect size/beta (\textbf{a,b}) and eQTL and GWAS significance/p-value (\textbf{c,d}) as a function of the GWAS class count.} \label{fig:beta_pval}
%%
%\end{figure}
